---
title: "FLash:::fwd for stock projection"
author: "Laurence Kell"
date: "August 13th, 2014"
output: rmarkdown::tufte_handout
---

# Introduction
\newthought{The precautionary approach}[^books_be] requires harvest control rules (HCRs) to trigger pre-agreed conservation and management action. This requires limit reference points to set boundaries that constrain harvesting within safe biological limits within which stocks can produce the maximum sustainable yield (MSY) and targets to ensure that management objectives are met. 

The performance of HCRs, i.e. how well they meet management objectives should be evaluated, ideally using Management Strategy Evaluation (MSE) where the HCRs is tested as part of a Management Procedure (MP). Where an MP is the combination of the data collection regime stock assessment procedure and the setting of managment regulations. HCRs can be modelled using the fwd method of FLR; see the MSE document for examples of simulation testing.

Simulating the evolution of a stock or population (i.e. a projection) may be required after an assessment for a range of catches to allow managers to decide upon a TAC or within an MSE for a management measure set by an MP.

fwd takes objects descibing historical stock status and assumptions about future dynamics(e.g. growth, maturity, natural mortality and recruitment dynamics), then performs a projection for future options e.g. for catches, fishing mortality. 


# Introduction
```{r}
library(FLCore)
library(FLash)
library(FLBRP)
library(ggplotFL)
```


fwd is used to make future projections, e.g. to evaluate different management options such as Total Allowable Catches (TACs) once a stock assessment has been conducted or for simulating a Harvest Control Rule (HCR) as part of a Management Strategy Evaluation (MSE).

In this examples we use the ple4 FLSock object

```{r}
data(ple4)
```

## The future
To perform a projection requires assumptions about future processes such as growth and recruitment and the effect of management on selectivity.

Recruit is based on a stock recruitment relationship, which can be fitted to the historic time series

```{r}
#### SRR
sr   =as.FLSR(ple4,model="bevholt")
sr   =fmle(sr,control=list(silent=TRUE))
```

While future growth and selectivity is often assummed to be an average of recent values. In which case these can be estimated using FLBRP. An advantage of using FLBRP is then the projections and reference points will be consistent.

```{r}
#### BRPs
eql=FLBRP(ple4,sr=sr)
computeRefpts(eql)
eql=brp(eql)

plot(eql)
```

Expected parameters in the future
```{r}
ggplot(FLQuants(eql,"m","mat","stock.wt","catch.sel"))+
  geom_line(aes(age,data))+
  facet_wrap(~qname,scale="free_y")
```

Setting up the projection years can then be done by  extending an FLStock uisng fwdWindow and passing the FLBRP object 
```{r}
stk=fwdWindow(ple4,end=2020,eql)
unlist(dims(stk))
```


## Projections

We first show how simple projections (e.g. for F and catch) can be performed. Later we show how a variety of HCRs can be simulated.

Simulate fishing at $F_{0.1}$, first create an FLQuant with the target Fs

```{r}
F0.1=FLQuant(refpts(eql)["f0.1","harvest",drop=T],dimnames=list(year=2009:2020))
```

Then project forward, note that sr is also required
```{r}
stk=fwd(stk,f=F0.1,sr=sr)

plot(FLStocks("Historic"=ple4,"F0.1"=stk))
```

It is possible to project for different Fs i.e. alternative reference points
```{r}
library(plyr)

dimnames(refpts(eql))$refpt

refs=refpts(eql)[c("msy","f0.1","fmax","spr.30"),"harvest",drop=T]

targetF=FLQuants(mlply(data.frame(refs),
                       function(refs) 
                           FLQuant(refs,dimnames=list(year=2009:2020))))

names(targetF)=names(refs)

names(targetF)[]="f"
stks=fwd(stk,targetF,sr=sr)

plot(stks)
```

or different multipliers of $F_{MSY}$
```{r}
msyTargets=FLQuants(mlply(seq(0,2,.25),
                       function(x) 
                           FLQuant(x*refs["msy"],dimnames=list(year=2009:2020))))

names(msyTargets)[]="f"

stks=fwd(stk,msyTargets,sr=sr)

plot(stks)
```

Catch projections are done in a similar way e.g. for $MSY$
```{r}
refpts(eql)["msy"]

refpts(eql)["msy",c("harvest","yield")]

msy=FLQuant(c(refpts(eql)["msy","yield"]),dimnames=list(year=2009:2020))

stks=fwd(stk,catch=msy,sr=sr)

plot(stks)
```

Compare F and Catch projections, e.g. for $MSY$ and $F_{MSY}$
```{r}
msys=FLQuants("f"    =targetF[[1]],
              "catch"=msy)
stks=fwd(stk,msys,sr=sr)

plot(stks)
```

If the projected catch is high you could simulate high Fs, however, there will be a cap of effort and capacity so in practice such high Fs may not be realised. Therefore there is a constraint on F.
```{r}
catch=FLQuant(c(refpts(eql)["msy","yield"])*2,
                          dimnames=list(year=2009:2020))

stk=fwd(stk,catch=catch,sr=sr)

plot(stk)
```
i.e. maxF, this allows an upper limit to be set on F

```{r}
stk=fwd(stk,catch=catch,sr=sr,maxF=1)

plot(stk)
```

This can also be used to model capacity
```{r}
capacity=FLQuant(1,dimnames=list(year=2009:2020))
q       =rlnorm(1,FLQuant(0,dimnames=list(year=2009:2020)),.2)
maxF    =q*capacity

stk=fwd(stk,catch=catch,sr=sr,maxF=maxF)

plot(stk)
```

Other quantities can be considered in projections as well as catch and F, i.e. ssb, biomass,  landings, discards, f, f.catch, f.landings, f.discards, effort, costs, revenue, profit, mnsz.

Management has two main options, i.e. setting effort (as in the examples above) or relative F-at-age by changing the selection pattern.
The selection pattern-at-age of landings is that of the catch less discards e.g.
```{r}
ggplot(FLQuants(eql,"catch.sel","discards.sel","landings.sel"))+
  geom_line(aes(age,data,col=qname))
```

In the FLStock object there are therefore 3 selection pattern components, and unfortunate three ways of calculating each. fwd uses computeCatch to re-estimate the catch.n, landings.n and discards.n before calculating future selection patterns. 
```{r}
catch(stk)<-computeCatch(stk)
```

In fwd the selection patterns are then calculated as harvest*discards.n/catch.n, harvest*landings.n/catch.n and discards.sel+landings.sel

Simulation of gears that get rid of discarding can be done by
```{r}
noDiscards=stk
discards.n(noDiscards)[,ac(2009:2020)]=0
catch.n(   noDiscards)[,ac(2009:2020)]<-landings.n(noDiscards)[,ac(2009:2020)]
catch(noDiscards)<-computeCatch(noDiscards)

## Note adjustment of harvest
harvest(noDiscards)[,ac(2009:2020)]=harvest(stk)[,ac(2009:2020)]*
  landings.n(stk)[,ac(2009:2020)]/catch.n(stk)[,ac(2009:2020)]

noDiscards=fwd(noDiscards,f=F0.1,sr=sr)
stk       =fwd(stk,       f=F0.1,sr=sr)

plot(FLStocks("No Discards"=noDiscards,"F0.1"=stk))
```


Non stationarity is seen in many biological processes, what happens if future fecundity decreases?
```{r}
poorFec=stk
mat(poorFec)[1:5,ac(2009:2020)]=c(0,0,0,0,.5)

poorFec=fwd(poorFec,f=F0.1,sr=sr)

plot(FLStocks("Reduced \nFecundity"=poorFec,"F0.1"=stk))
```

# Stochasticity

Monte Carlo simulations based on future recruitment
```{r}
srDev=rlnorm(100,FLQuant(0,dimnames=list(year=2009:2020)),0.3)
plot(srDev)
```

```{r}
stk  =fwd(stk,f=F0.1,sr=sr,sr.residuals=srDev)
```


# fwdControl

fwdControl is a more flexible but fiddly way of setting up projections. For example to replicate the $F_{0.1}$ projection above requires setting up a fwdControl object.

This can be done using a constructor and a data.frame

```{r}
ctrl=fwdControl(data.frame(year    =2009:2018,
                           val     =c(refpts(eql)["f0.1","harvest"]),
                           quantity="f"))
```

fwdControl is a class with 5 slots

```{r}
slotNames(ctrl)
```

For now we will concerntrate on just the target and trgtArray slots.

```{r}
slotNames(ctrl)
ctrl
```

target specifies the quantity for the projection (e.g. "f", "catch", "ssb", ...) and the projection year. The projection can be a target by specifying it in val. While min and max specify bounds. For example if you want to project for a target F but also to check that SSB does not fall below an SSB limit. 

An example with high F that decreases SSB a lot

```{r}
target=fwdControl(data.frame(year=2009,val=0.8,quantity="f")) 
stk   =fwdWindow(ple4,end=2010,eql)

stk=fwd(stk,ctrl=target,sr=eql)

fbar(stk)[,"2009"]
ssb( stk)[,"2010"]
```

Note that it is the end of year biomass that is constrained as in this case spawning is at Jan 1st and so fishing only has an effect of SSB next year

Constrain SSB so that it doesnt fall below 250000
```{r}
target  <-fwdControl(data.frame(year=c(2009,2009),
                             val =c( 0.8,  NA),
                             min =c(NA,230000),
                             quantity=c("f","ssb")))

stk=fwd(stk,ctrl=target,sr=sr)

fbar(stk)[,"2009"]
ssb( stk)[,"2010"]
```


If a stock spawns mid year so the adult population is affected by fishing then the SSB constraint is within year, e.g.
```{r}
harvest.spwn(stk)[]=0.5

stk=fwd(stk,ctrl=target,sr=sr)

fbar(stk)[,  "2009"]
ssb( stk)[,c("2009","2010")]
```


```{r}
albF1     =fwd(stk,ctrl=target,sr=sr)
plot(albF1)
```


```{r}
ctrl      =fwdControl(data.frame(year    =2009:2020,
                                 val     =c(refpts(eql)["f0.1","harvest"])*0.5,
                                 quantity="f"))
stk=fwdWindow(ple4,end=2020,eql)
albF2     =fwd(stk, ctrl=ctrl, sr=sr)
```


```{r}
ctrl     =fwdControl(data.frame(year    =2009:2020,
                                val     =c(refpts(eql)["f0.1","harvest"])*2.0,
                                quantity="f"))
albF3     =fwd(stk, ctrl=ctrl, sr=sr)
```


```{r,eval=FALSE}
## Create an FlStock object
albF0.1=FLStocks("F0.1"=albF1,"half"=albF2,"double"=albF3)
plot(albF0.1)

## Cut the plots
plot(lapply(albF0.1,window,start=1990))

## Compare alternatives
lapply(lapply(albF0.1,window,start=2009),computeCatch)

#### Total catch
lapply(lapply(lapply(albF0.1,window,start=2009),computeCatch),sum)

#### Short-term
unlist(lapply(lapply(lapply(albF0.1,window,start=2009,end=2014),computeCatch),sum))

#### Medium-term
unlist(lapply(lapply(lapply(albF0.1,window,start=2017,end=2021),computeCatch),sum))

#### Long-term
unlist(lapply(lapply(lapply(albF0.1,window,start=2024,end=2028),computeCatch),sum))
```


```{r1,eval=FALSE}
#### constant catch startegies
ctch=mean(computeCatch(albNEA)[,ac(2004:2008)])

albC=FLStocks()
ctrl=fwdControl(data.frame(year=2009:2028,val=ctch,quantity="catch"))
albC[["1.0"]]     =fwd(stk,ctrl=ctrl,sr=sr)

ctrl     =fwdControl(data.frame(year=2009:2028,val=0.5*ctch,quantity="catch"))
albC[["0.5"]]     =fwd(stk,ctrl=ctrl,sr=sr)

ctrl     =fwdControl(data.frame(year=2009:2028,val=1.5*ctch,quantity="catch"))
albC[["1.5"]]     =fwd(stk,ctrl=ctrl,sr=sr)
plot(albC)

#### compare strategies
plot(FLStocks(albC[[1]],albF0.1[[1]]))
```


```{r2,eval=FALSE}
#### constant catch with an upper F bound
ctrl=fwdControl(data.frame(year    =rep(2009:2028,each=20),
                            val     =rep(c(ctch*1.5,NA),20),
  		      max     =rep(c(NA,F0.1),20),
                            quantity=rep(c("catch","f"),20)))
albFC=fwd(stk,ctrl=ctrl,sr=sr)
```


```{r3,eval=FALSE}
#### 5% F reduction
ctrl=fwdControl(data.frame(year    =rep(2009:2028,each=2),
                            rel.year=c(t(array(c(2008:2027,rep(NA,20)),c(20,2)))),
                            val     =rep(c(0.95,NA),20),
                            min     =rep(c(NA,F0.1*.5),20),
                            quantity=rep(c("catch","f"),20)))
albFC=fwd(stk,ctrl=ctrl,sr=sr)
plot(albFC)
```


```{r4,eval=FALSE}
#### 10% SSB increase 
ctrl=fwdControl(data.frame(year    =2009:2028,
                            rel.year=2008:2027,
                            min     =1.10,
                            quantity="ssb"))
albSSB=fwd(stk,ctrl=ctrl,sr=sr)
plot(albSSB)
```



```{r5,eval=FALSE}
hcrF=function(iYr,SSB,Bpa,Blim,Fmin,Fmax){
    val =pmin(Fmax,Fmax-(Fmax-Fmin)*(Bpa-SSB)/(Bpa-Blim))
    trgt=fwdTarget(year=iYr+1,quantity="f",valueval)

    return(trgt)}
```


```{r6,eval=FALSE}
data(ple4)

# Set up the stock for the next 6 years
pleProj =stf(ple4,6)

# Set a constant recruitment based on the geometric mean of last 10 years
mnRec = FLPar(exp(mean(log(rec(ple4)[,ac(1992:2001)]))))
# Set ssb target to level 19 years ago
ssbTarget = ssb(ple4)[,"1992"]

## function to minimise
f = function(x,stk,ssbTarget,ctrl,sr)
       {
       ctrl@target[,"val"]    =x
       ctrl@trgtArray[,"val",]=x

       ssb.=c(ssb(fwd(stk,ctrl=ctrl,sr=sr))[,"2006"])

       return((ssb.-ssbTarget)^2)
       }

## Recover stock to BMY in 2006 with a constant F strategy
ctrl=fwdControl(data.frame(year=2002:2006,val=.5,rel=2001,quantity="f"))

xmin=optimize(f, c(0.1, 1.0), tol = 0.0000001, stk=pleProj, ssbTarget=ssbTarget, ctrl=ctrl, sr=list(model="mean",params=mnRec))
ctrl=fwdControl(data.frame(year=2002:2006,val=xmin$minimum,rel=2001,quantity="f"))
pleProjF     =fwd(pleProj,ctrl=ctrl,sr=list(model="mean", params=mnRec))

# update catch slot
catch(pleProjF) = computeCatch(pleProjF)

# Have we reached the target?
ssbTarget
ssb(pleProjF)[,ac(2002:2006)]
# At what level of constant F
fbar(pleProjF)[,ac(2002:2006)]
# 'ave a butchers
plot(pleProjF[,ac(1957:2006)])

plot(albSSB)
```


```{r7,eval=FALSE}

data(ple4)
pleProj=stf(ple4,6)

## Recover stock to the desired SSB in 2006 with a constant Catch strategy
# Here val can be anything in the ctrl because it is overwritten in the optimisation loop
ctrl=fwdControl(data.frame(year=2002:2006,val=c(catch(pleProj)[,"2001"]),quantity="catch"))

xmin=optimize(f, c(100, 100000), tol = 0.0000001, stk=pleProj, ssbTarget=ssbTarget, ctrl=ctrl, sr=list(model="mean",params=mnRec))
ctrl=fwdControl(data.frame(year=2002:2006,val=xmin$minimum,quantity="catch"))
pleProjC      =fwd(pleProj,ctrl=ctrl,sr=list(model="mean", params=mnRec))

# Have we reached the target?
ssbTarget
ssb(pleProjC)[,ac(2002:2006)]
# At what level of constant catch
computeCatch(pleProjC)[,ac(2002:2006)]
# And at what level of F
fbar(pleProjC)[,ac(2002:2006)]
# Update the catch slot
catch(pleProjC) = computeCatch(pleProjC)
# 'ave a butchers
plot(pleProjC[,ac(1957:2006)])
```


```{r8,eval=FALSE}
# Assessment up to and including 2001
data(ple4)
black.bird               =stf(pleProj,nyrs=2)

# set courtship and egg laying in Autumn
black.bird@m.spwn[]      =0.66
black.bird@harvest.spwn[]=0.66

# assessment is in year 2002, set catch constraint in 2002 and a first guess for F in 2003
ctrl          =fwdControl(data.frame(year=2002:2003,val=c(85000,.5),quantity=c("catch","f")))
black.bird    =fwd(black.bird, ctrl=ctrl, sr=list(model="mean", params=FLPar(25000)))

# HCR specifies F=0.1 if ssb<100000, F=0.5 if ssb>300000
# otherwise linear increase as SSB increases
min.ssb=100000
max.ssb=300000
min.f  =0.1
max.f  =0.5

# slope of HCR
a.    =(max.f-min.f)/(max.ssb-min.ssb)
b.    =min.f-a.*min.ssb

# plot of HCR
plot(c(0.0,min.ssb,max.ssb,max.ssb*2),c(min.f,min.f,max.f,max.f),type="l",ylim=c(0,max.f*1.25),xlim=c(0,max.ssb*2))

## find F through iteration
t.    =999
i     =0
while (abs(ctrl@target[2,"val"]-t.)>10e-6 & i<50)
   {
   t.=ctrl@target[2,"val"]  ## save last val of F

   # calculate new F based on SSB last iter
   ctrl@target[2,"val"]    =a.*c(ssb(black.bird)[,"2003"])+b.
   ctrl@trgtArray[2,"val",]=a.*c(ssb(black.bird)[,"2003"])+b.
   black.bird=fwd(black.bird, ctrl=ctrl, sr=list(model="mean", params=FLPar(25000)))

   # 'av a gander
   points(c(ssb(black.bird)[,"2003"]),c(ctrl@target[2,"val"]),cex=1.25,pch=19,col=i)
   print(c(ssb(black.bird)[,"2003"]))
   print(c(ctrl@target[2,"val"]))
   i=i+1
   }

# F bounds
black.bird      =fwd(black.bird, ctrl=ctrl, sr=list(model="mean",params=FLPar(25000)))
plot(FLStocks(black.bird))
```


```{r9,eval=FALSE}
#### Create a random variable for M
albM   =stk
m(albM)=propagate(m(albM),100)

mDev=rlnorm(prod(dim(m(albM))),0,0.3)
mean(mDev)
var(mDev)^.5

m(albM)=m(albM)*FLQuant(mDev,dimnames=dimnames(m(albM)))
plot(m(albM))

harvest(albM)=computeHarvest(albM)
catch(  albM)=computeCatch(  albM,"all")

plot(FLStocks(albM,stk28))
 
ctrl=fwdControl(data.frame(year=2009:2028,val=ctch,quantity="catch"))
albM     =fwd(albM,ctrl=ctrl,sr=sr)

plot(albM)
```

```{r10,eval=FALSE}
#### Create a random variable for M
albM1   =albM
m(albM1)[1:3,]          =m(albM)[1:3,]*2

harvest(albM1)=computeHarvest(albM1)
catch(  albM1)=computeCatch(  albM1,"all")
albM1         =fwd(albM1,ctrl=ctrl,sr=sr)

albM2   =albM
m(albM2)[,ac(2000:2028)]=m(albM)[,ac(2000:2028)]*2

harvest(albM2)=computeHarvest(albM2)
catch(  albM2)=computeCatch(  albM2,"all")
albM2         =fwd(albM2,ctrl=ctrl,sr=sr)
 
plot(FLStocks(albM,albM1,albM2))
```

```{r11,eval=FALSE}
#### process error in recruitment
srDev=FLQuant(rlnorm(20*100,0.0,0.3),dimnames=list(year=2008:2028,iter=1:100))
sr=fwd(albM,ctrl=ctrl,sr=sr,sr.residuals=srDev)
plot(sr)
```

```{r12,eval=FALSE}
#### SRR regime shifts
albSV       =as.FLSR(albNEA)
model(sr)=bevholtSV()
albSV =fmle(albSV,fixed=list(spr0(albNEA)))
albSV1=fmle(albSV,fixed=list(spr0=spr0(albNEA),s=0.75))
albSV2=fmle(albSV,fixed=list(spr0=spr0(albNEA),v=0.75*params(albSV)["v"]))

#### Prior for steepness
albSV3=fmle(albSV,fixed=list(s=qnorm(seq(0.01,0.99,length.out=101),.75,.1)))
albSV3=fwd(albSV3,ctrl=ctrl,sr=sr,sr.residuals=srDev)

plot(albSV1,albSV2,albSV3)
```

```{r13,eval=FALSE}
#### SRR regime shifts
albBRP=brp(FLBRP(albM))
refpts(albBRP)

albSV3=fmle(albSV,fixed=list(s=qnorm(seq(0.01,0.99,length.out=101),.75,.1)))
albSV3=fwd(albSV3,ctrl=ctrl,sr=sr,sr.residuals=srDev)

plot(albSV1,albSV2,albSV3)
```

```{r14,eval=FALSE}
# F bounds
black.bird      =fwd(black.bird, ctrl=ctrl, sr=list(model="mean",params=FLPar(25000)))
plot(FLStocks(black.bird))
```

```{r15,eval=FALSE}
library(FLash)
library(FLAssess)

#### Set up a short term forecast for an FLStock object by adding extra years
## The default forecast is 3 years,
alb3=stf(alb)

## Check what?s happened
summary(alb)
summary(alb3)

## by default future F is the mean of last 3 years
mean(fbar(alb)[,ac(2007-(0:2))])
fbar(alb3)[,ac(2007+(1:3))]

## by default future F is the mean of last 3 years
mean(fbar(alb)[,ac(2007-(0:2))])
fbar(alb3)[,ac(2007+(1:3))]
```

```{r16,eval=FALSE}
## Constant F Projection for a 20 year projection
stk=stf(alb,nyear=20)

#### SRR
sr       =as.FLSR(alb)
model(sr)=bevholt()
sr       =fmle(sr)

#### BRPs
albBRP=FLBRP(alb,sr=sr)
computeRefpts(albBRP)


albBRP=brp(albBRP)


# Use F0.1 as fishing mortality target
F0.1=refpts(albBRP)["f0.1","harvest",drop=T]
#### bug
ctrl     =fwdControl(data.frame(
               year    =2008:2027,
               val     =F0.1,
               quantity="f"))

albF1     =fwd(stk,ctrl=ctrl,sr=sr)

plot(albF1)
ctrl     =fwdControl(data.frame(
               year    =2008:2027,
               val     =F0.1*0.5,
               quantity="f"))
albF2     =fwd(stk, ctrl=ctrl, sr=sr)

ctrl     =fwdControl(data.frame(
               year    =2008:2027,
               val     =F0.1*2.0,
               quantity="f"))
albF3     =fwd(stk, ctrl=ctrl, sr=sr)


## Create an FlStock object
albF0.1=FLStocks("F0.1"=albF1,"half"=albF2,"double"=albF3)
plot(albF0.1)

## Cut the plots
plot(lapply(albF0.1,window,start=1990))

## Compare alternatives
lapply(lapply(albF0.1,window,start=2008),computeCatch)

#### Total catch
lapply(lapply(lapply(albF0.1,window,start=2008),computeCatch),sum)

#### Short-term
unlist(lapply(lapply(lapply(albF0.1,window,start=2008,end=2013),computeCatch),sum))
#### Medium-term
unlist(lapply(lapply(lapply(albF0.1,window,start=2016,end=2020),computeCatch),sum))
#### Long-term
unlist(lapply(lapply(lapply(albF0.1,window,start=2023,end=2027),computeCatch),sum))

```

```{r17,eval=FALSE}
#### constant catch startegies
ctch=mean(computeCatch(alb)[,ac(2003:2007)])

albC=FLStocks()
ctrl=fwdControl(data.frame(year=2008:2027,val=ctch,quantity="catch"))
albC[["1.0"]]     =fwd(stk,ctrl=ctrl,sr=sr)

ctrl     =fwdControl(data.frame(year=2008:2027,val=0.5*ctch,quantity="catch"))
albC[["0.5"]]     =fwd(stk,ctrl=ctrl,sr=sr)

ctrl     =fwdControl(data.frame(year=2008:2027,val=1.5*ctch,quantity="catch"))
albC[["1.5"]]     =fwd(stk,ctrl=ctrl,sr=sr)
plot(albC)

#### compare startegies
plot(FLStocks(albC[[1]],albF0.1[[1]]))

```

```{r18,eval=FALSE}
#### constant catch with an upper F bound
ctrl=fwdControl(data.frame(year    =rep(2008:2027,each=20),
                            val     =rep(c(ctch*1.5,NA),20),
  		      max     =rep(c(NA,F0.1),20),
                            quantity=rep(c("catch","f"),20)))
albFC=fwd(stk,ctrl=ctrl,sr=sr)
plot(albFC)
```

```{r19,eval=FALSE}
#### 5% F reduction
ctrl=fwdControl(data.frame(year    =rep(2008:2027,each=2),
                            rel.year=c(t(array(c(2007:2026,rep(NA,20)),c(20,2)))),
                            val     =rep(c(0.95,NA),20),
                            min     =rep(c(NA,F0.1*.5),20),
                            quantity=rep(c("catch","f"),20)))
albFC=fwd(stk,ctrl=ctrl,sr=sr)
plot(albFC)
```

```{r20,eval=FALSE}
#### 10% SSB increase
ctrl=fwdControl(data.frame(year    =2008:2027,
                            rel.year=2007:2026,
                            min     =1.10,
                            quantity="ssb"))
albSSB=fwd(stk,ctrl=ctrl,sr=sr)
plot(albSSB)
```

```{r21,eval=FALSE}
hcrF=function(iYr,SSB,Bpa,Blim,Fmin,Fmax){
    val =pmin(Fmax,Fmax-(Fmax-Fmin)*(Bpa-SSB)/(Bpa-Blim))
    trgt=fwdTarget(year=iYr+1,quantity="f",valueval)

    return(trgt)}
```

```{r22,eval=FALSE}
## Ogives
dnormal=function(x,a,sL,sR){
  pow=function(a,b) a^b

  func=function(x,a,sL,sR){
    if (x < a) return(pow(2.0,-((x-a)/sL*(x-a)/sL)))
    else       return(pow(2.0,-((x-a)/sR*(x-a)/sR)))}

  sapply(x,func,a,sL,sR)}

logistic=function(x,a50,ato95){
  pow=function(a,b) a^b

  func=function(x,a50,ato95){
     if ((a50-x)/ato95 > 5)   return(0)
     if ((a50-x)/ato95 < -5)  return(1)

     return(1.0/(1.0+pow(19.0,(a50-x)/ato95)))}

  sapply(x,func,a50,ato95)}

prices    =data.frame(rbind(cbind(Age=1:10,Price=dnormal( 1:10,3,10,20),Type="Peaking"),
                             cbind(age=1:10,Price=logistic(1:10,2,3),    Type="Increasing")))
prices$Age=as.numeric(ac(prices$Age))

p    = ggplot(prices,aes(x=Age, y=Price, group=Type))
p    = p + geom_line(aes(colour=Type))
p

refIPrice=brp(FLBRP(alb,fbar=seq(0,1,length.out=101)))
refPPrice=refIPrice

price(refIPrice)=logistic(1:15,4,3)
price(refPPrice)=dnormal( 1:15,5,1,5)

refIPrice=brp(refIPrice)
refPPrice=brp(refPPrice)

breakEven=refIPrice
#### bug why not no recycling
refpts(breakEven)=refpts(as.numeric(c(refpts(refIPrice)["fmax","revenue"]*2,rep(NA,7))),refpt=c("breakEven"))
computeRefpts(breakEven)[,"revenue"]

vcost(refIPrice)=c(computeRefpts(breakEven)[,"revenue"]*0.20)
fcost(refIPrice)=vcost(refIPrice)*4.0

vcost(refPPrice)=vcost(refIPrice)
fcost(refPPrice)=fcost(refIPrice)

refIPrice=brp(refIPrice)
refPPrice=brp(refPPrice)

price(refIPrice)=price(refIPrice)/c(refpts(refIPrice)["mey","profit"])
price(refPPrice)=price(refPPrice)/c(refpts(refPPrice)["mey","profit"])

refIPrice=brp(refIPrice)
refPPrice=brp(refPPrice)

plot(refPPrice)
plot(refIPrice)
```

```{r23,eval=FALSE}
data(ple4)

# Set up the stock for the next 6 years
pleProj =stf(ple4,6)

# Set a constant recruitment based on the geometric mean of last 10 years
mnRec = FLPar(exp(mean(log(rec(ple4)[,ac(1992:2001)]))))
# Set ssb target to level 19 years ago
ssbTarget = ssb(ple4)[,"1992"]

## function to minimise
f = function(x,stk,ssbTarget,ctrl,sr)
       {
       ctrl@target[,"val"]    =x
       ctrl@trgtArray[,"val",]=x

       ssb.=c(ssb(fwd(stk,ctrl=ctrl,sr=sr))[,"2006"])

       return((ssb.-ssbTarget)^2)
       }

## Recover stock to BMY in 2006 with a constant F strategy
ctrl=fwdControl(data.frame(year=2002:2006,val=.5,rel=2001,quantity="f"))

xmin=optimize(f, c(0.1, 1.0), tol = 0.0000001, stk=pleProj, ssbTarget=ssbTarget, ctrl=ctrl, sr=list(model="mean",params=mnRec))
ctrl=fwdControl(data.frame(year=2002:2006,val=xmin$minimum,rel=2001,quantity="f"))
pleProjF     =fwd(pleProj,ctrl=ctrl,sr=list(model="mean", params=mnRec))

# update catch slot
catch(pleProjF) = computeCatch(pleProjF)

# Have we reached the target?
ssbTarget
ssb(pleProjF)[,ac(2002:2006)]
# At what level of constant F
fbar(pleProjF)[,ac(2002:2006)]
# 'ave a butchers
plot(pleProjF[,ac(1957:2006)])

plot(albSSB)
```

```{r24,eval=FALSE}
data(ple4)
pleProj=stf(ple4,6)

## Recover stock to the desired SSB in 2006 with a constant Catch strategy
# Here val can be anything in the ctrl because it is overwritten in the optimisation loop
ctrl=fwdControl(data.frame(year=2002:2006,val=c(catch(pleProj)[,"2001"]),quantity="catch"))

xmin=optimize(f, c(100, 100000), tol = 0.0000001, stk=pleProj, ssbTarget=ssbTarget, ctrl=ctrl, sr=list(model="mean",params=mnRec))
ctrl=fwdControl(data.frame(year=2002:2006,val=xmin$minimum,quantity="catch"))
pleProjC      =fwd(pleProj,ctrl=ctrl,sr=list(model="mean", params=mnRec))

# Have we reached the target?
ssbTarget
ssb(pleProjC)[,ac(2002:2006)]
# At what level of constant catch
computeCatch(pleProjC)[,ac(2002:2006)]
# And at what level of F
fbar(pleProjC)[,ac(2002:2006)]
# Update the catch slot
catch(pleProjC) = computeCatch(pleProjC)

plot(pleProjC[,ac(1957:2006)])
```


```{r25,eval=FALSE}
# Assessment upto and including 2001
data(ple4)
black.bird               =stf(pleProj,nyear=2)

# set courtship and egg laying in Autumn
black.bird@m.spwn[]      =0.66
black.bird@harvest.spwn[]=0.66

# assessment is in year 2002, set catch constraint in 2002 and a first guess for F in 2003
ctrl          =fwdControl(data.frame(year=2002:2003,val=c(85000,.5),quantity=c("catch","f")))
black.bird    =fwd(black.bird, ctrl=ctrl, sr=list(model="mean", params=FLPar(25000)))

# HCR specifies F=0.1 if ssb<100000, F=0.5 if ssb>300000
# otherwise linear increase as SSB increases
min.ssb=100000
max.ssb=300000
min.f  =0.1
max.f  =0.5

# slope of HCR
a.    =(max.f-min.f)/(max.ssb-min.ssb)
b.    =min.f-a.*min.ssb

# plot of HCR
plot(c(0.0,min.ssb,max.ssb,max.ssb*2),c(min.f,min.f,max.f,max.f),type="l",ylim=c(0,max.f*1.25),xlim=c(0,max.ssb*2))

## find F through iteration
t.    =999
i     =0
while (abs(ctrl@target[2,"val"]-t.)>10e-6 & i<50)
   {
   t.=ctrl@target[2,"val"]  ## save last val of F

   # calculate new F based on SSB last iter
   ctrl@target[2,"val"]    =a.*c(ssb(black.bird)[,"2003"])+b.
   ctrl@trgtArray[2,"val",]=a.*c(ssb(black.bird)[,"2003"])+b.
   black.bird=fwd(black.bird, ctrl=ctrl, sr=list(model="mean", params=FLPar(25000)))

   # 'av a gander
   points(c(ssb(black.bird)[,"2003"]),c(ctrl@target[2,"val"]),cex=1.25,pch=19,col=i)
   print(c(ssb(black.bird)[,"2003"]))
   print(c(ctrl@target[2,"val"]))
   i=i+1
   }

# F bounds
black.bird      =fwd(black.bird, ctrl=ctrl, sr=list(model="mean",params=FLPar(25000)))
plot(FLStocks(black.bird))
```

```{r26,eval=FALSE}
#### Create a random variable for M
albM   =albF1
m(albM)=propagate(m(albM),100)

mDev=rlnorm(prod(dim(m(albM))),0,0.3)
mean(mDev)
var(mDev)^.5

m(albM)=m(albM)*FLQuant(mDev,dimnames=dimnames(m(albM)))
plot(m(albM))

harvest(albM)=computeHarvest(albM)
catch(  albM)=computeCatch(  albM,"all")

ctrl=fwdControl(data.frame(year=2008:2027, val=c(fbar(albF1)[,ac(2008:2027)]),quantity="f"))
albM=fwd(albM,ctrl=ctrl,sr=sr)

plot(FLStocks(albM,albF1))
```

```{r27,eval=FALSE}
#### Create a random variable for M
albM1	   =albM
m(albM1)[1:3,]          =m(albM)[1:3,]*2

harvest(albM1)=computeHarvest(albM1)
catch(  albM1)=computeCatch(  albM1,"all")
albM1         =fwd(albM1,ctrl=ctrl,sr=sr)

albM2   =albM
m(albM2)[,ac(2000:2027)]=m(albM)[,ac(2000:2027)]*2

harvest(albM2)=computeHarvest(albM2)
catch(  albM2)=computeCatch(  albM2,"all")
albM2         =fwd(albM2,ctrl=ctrl,sr=sr)

plot(FLStocks(albM,albM1,albM2))
```

```{r28,eval=FALSE}
#### process error in recruitment
srDev=FLQuant(rlnorm(20*100,0.0,0.3),dimnames=list(year=2008:2027,iter=1:100))
sr=fwd(albM,ctrl=ctrl,sr=sr,sr.residuals=srDev)
plot(sr)
```


```{r29,eval=FALSE}
sr =as.FLSR(alb,model="bevholtSV")
sr1=fmle(sr,fixed=list(spr0=spr0(alb)))

#### SRR regime shifts
sr2=fmle(sr,fixed=list(spr0=spr0(alb),v=0.75*params(sr)["v"]))

alb2=fwd( sr3,ctrl=ctrl,sr=sr2,sr.residuals=srDev)

plot(FLStocks(sr,sr2))
```


# Syntax

## Biology
## Selectivity
## Monte Carlo simulation  

# Targets  
## Catch and F strategies	

# Limits 	
## Relative targets and limits	

# Harvest Control Rules	

# Multi-annual management  

# Recovery Plans	
# Long-term plans 	
# Technical measures	


# fwdControl  
## target   
## trgtArray   
## effort 	
## effArray 	
## blocks	

# Management Objectives  
# Performance statistics	

# Examples	
## Albacore	
## Herring	
## Cod	
## North Sea flatfish	
## Bluefin 	
## Mediterannean Swordfish	

# References	

# fwd  
